---
title: "UNICEF Child Labour & Life Expectancy Analysis"
toc: TRUE
format: 
  html:
    self-contained: true
editor: visual
---


```{r, warning=FALSE, message=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(sf)
library(leaflet)
```

## Gender Wise Child Labour:

A bar chart has been constructed to represent the gender-specific child labour threshold and is attached below. Surprisingly, the child labour barrier is higher for females than for males, at 66.58% and 55.27%, respectively. The total child labour threshold is 60.77%, which is quite high.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
data <- read.csv("unicef_indicator_1.csv")
data2 <- read.csv("unicef_metadata.csv")

master <- data %>% inner_join(data2, by = "country")

sexWise_ChildLabour <- master %>%
  group_by(sex) %>%
  summarise(Avg_Value = round(mean(obs_value, na.rm = T), 2)) %>%
  arrange(-Avg_Value)

# Create the bar chart
ggplot(sexWise_ChildLabour, aes(x = sex, y = Avg_Value, fill = sex)) +
  geom_bar(stat = "identity", width = 0.7) +
  geom_text(aes(label = Avg_Value),
            vjust = 1.0, color = "white", size = 3) +
  labs(title = "Gender Wise Average Child Labour Thresholds", 
       x = "Gender",
       y = "Average Child Labour Threshold") +
  guides(fill = FALSE) 
```

## Relationship Between Military Expense of GDP and Life Expectancy:

A scatter chart is generated to see whether there is a relationship between military expenditure on GDP and life expectancy at birth. A scatter plot with a regression line fitted to it is created and attached below. The regression line is downward, indicating a negative association between the two variables. This suggests that increasing military expenditure of GDP reduces life expectancy at birth, but decreasing military expenditure of GDP increases life expectancy at birth.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# Create the scatter plot with regression line
ggplot(master, aes(x = Military.expenditure....of.GDP., 
                   y = Life.expectancy.at.birth..total..years.)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship Between Military Expenditure & Life Expectancy",
       x = "Military Expenditure of GDP", y = "Life Expectancy")+
  theme_minimal()
```

## Trend of Military Expenditure of GDP:

A bar chart was developed to depict the average military expenditure of GDP over time, which is attached below. It is clear that the average military expenditure as a percentage of GDP has been decreasing over time. We notice a very significant negative trend, indicating that average life expectancy increases over time. After 1982, the military expense of GDP decreased.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
ME_Year <- master %>%
  group_by(year) %>%
  summarise(Military_Expenditure = mean(Military.expenditure....of.GDP.,
                                   na.rm = T))

# Line chart with markers
ggplot(ME_Year, aes(x = year, y = Military_Expenditure, group = 1)) +
  geom_line(lwd = 1) +
  geom_point(color = "red", size = 2) + 
  labs(title = "Average Military Expenditure of GDP Over Time", 
       x = "Year", 
       y = "Military Expenditure of GDP") +
  theme_minimal()
```

## Military Expenditure of GDP By Country:

Finally, a world map is created which is attached below. The map shows the average military expenditure of GDP by each country. The top 3 countries with highest average military expenditure of GDP are Israel, Saudia Arabia and Qatar.

```{r, warning=FALSE, message=FALSE, echo=FALSE}
# Calculate mean military expenditure by country
ME_Country <- master %>%
  group_by(alpha_3_code.x) %>%
  summarise(Military_Expenditure = mean(Military.expenditure....of.GDP.,
                                        na.rm = TRUE))
names(ME_Country)[1] <- "ISO3_CODE"

# Load country polygons and get centroids
final <- giscoR::gisco_get_countries() %>% 
  st_centroid() %>%
  inner_join(ME_Country, by = "ISO3_CODE")

# Extract latitude and longitude from the 'geometry' column
final$latitude <- st_coordinates(final)[, 2]
final$longitude <- st_coordinates(final)[, 1]

# Define a function to scale the radius based on military expenditure
scale_radius <- function(x) {
  # Define the range of circle marker sizes
  min_radius <- 3
  max_radius <- 15
  
  # Scale the military expenditure values to the range of circle marker sizes
  scaled_values <- scales::rescale(x, to = c(min_radius, max_radius))
  
  return(scaled_values)
}

# Create a leaflet map
m_leaflet <- leaflet() %>%
  addTiles() %>%
  setView(lng = 0, lat = 30, zoom = 2)

# Add markers for each country with military expenditure as popup
m_leaflet %>% 
  addCircleMarkers(
    data = final,
    lng = ~longitude,  
    lat = ~latitude,
    radius = ~scale_radius(Military_Expenditure),  # Use scaled radius based on military expenditure
    popup = ~paste(NAME_ENGL, "<br>", "Military Expenditure:", round(Military_Expenditure, 2)),
    label = ~NAME_ENGL,
    color = "red",
    fillOpacity = 0.8
  )
```


